Carbonate chemistry at Ny-Ålesund: AWIPEV-CO~2~ Project

### 2020-08-28 awipev-CO2.Rmd
still on range of seafet pH

```{r Read, cleanup and prepare ferrybox data, echo=FALSE, warning=FALSE, message=FALSE}
d_hour <-
  readRDS(file = paste0(
    path,
    "fb_awipev-co2_server/ny-alesund/data/NRT_data/nydata_hour.rds"
  ))
d_hour <- d_hour %>%
  dplyr::rename(
    pco2_corr = pco2_corr_filtered,
    temp_dur = temp_dur_filtered,
    pco2_raw = pco2_raw_filtered,
    sal_fb = sal_fb_filtered,
    temp_insitu_11m = temp_insitu_11m_filtered,
    temp_fb = temp_fb_filtered,
    ph_dur = ph_dur_filtered
  ) %>% 
  arrange(datetime)
d_all <- d_hour #d_all will be the merger
d_all <- d_all %>%
  dplyr::mutate_all(function(x) ifelse(is.nan(x), NA, x)) %>% # replace all NaN by NA
  dplyr::rowwise() %>% 
  dplyr::mutate(n_depth_ctd = sum(is.na( # create new variable to identify the hours for which there are CTD data at more than 1 depth 
  c(
    sal_insitu_ctd_1m,
    sal_insitu_ctd_3m,
    sal_insitu_ctd_5m,
    sal_insitu_ctd_7m,
    sal_insitu_ctd_9m
  )
), na.rm = TRUE)) %>% 
  ungroup()
#restore datetime, date which are mangled by the mutates above, arrange ascending
d_all <- d_all %>% 
  dplyr::mutate(datetime = d_hour$datetime, 
                       date = d_hour$date,
                       month = month(d_hour$datetime))
```

@@

# seaFET pH 
discrete2merge <- discrete %>%
  dplyr::select(closest_datetime, ph_s_sf_temp_insi) %>%
  dplyr::filter(!is.na(closest_datetime)) %>%
  rename(datetime=closest_datetime)
d_all <- full_join(d_all, discrete2merge, by="datetime")
d_all <- d_all %>% 
  dplyr::mutate(ph_s_sf_temp_insi = ifelse(n_depth_ctd == 4, ph_s_sf_temp_insi, NA)) %>% # eliminate ph spec data when the profiler was at several depth during the same hour
  dplyr::mutate(ph_s_sf_temp_insi = ifelse(!is.na(temp_insitu_ctd_7m) |
                                             !is.na(temp_insitu_ctd_9m), 
                                           ph_s_sf_temp_insi, NA)) # only use ph spec data when profiler was below 6 m

@@

# bind new calibrated pH data to the seafet data set.
names(seafet_calib) <- c("pHint_tot_sf", "pHext_tot_sf")
d_all <- cbind(d_all, seafet_calib )

# pH seaFET Final
# add despike with n=2 k = 5762 to the final ph corrected data
# despike me semble inutile, faire plutôt un simple des outliers qui, vérification faite, n'est même pas nécessaire
# Il faut par contre créér une nouvelle variable avec pH deeper than 8 m, called *_9m for consistency with salinity and temp
d_all <- d_all %>%     
  #dplyr::mutate(pHint_tot_sf = despike(d_all$pHint_tot_sf, reference= "median", n=2, k=5761, replace="NA"))
  tt <- dplyr::mutate(tmp, pHint_tot_sf_9m = ifelse(pressure_insitu_ctd > 8, pHint_tot_sf, NA))

@@

p <- ggplot(d_all, aes(x=datetime, y= pHint_tot_sf))+
  geom_point(size=1, alpha=0.5, na.rm=TRUE)

seafet_depth_xts <- dplyr::select(d_all, datetime, ph_s_sf_temp_insi, pHint_tot_sf, pH_s_sf_no_calb)
seafet_depth_xts <- as.xts(seafet_depth_xts, order.by = d_all$datetime)
dygraph(seafet_depth_xts, group = "depth", main="pH seaFET (ref + corrected + raw)", ylab="pHT") %>%
dySeries("pHint_tot_sf", label = "pHT corrected", color =  RColorBrewer::brewer.pal(3, "Dark2")[1], strokeWidth = 0, pointSize = 3) %>%
dySeries("ph_s_sf_temp_insi", label = "pH spectro for calibration", color =  "red", strokeWidth = 0, pointSize = 4 ) %>%
dySeries("pH_s_sf_no_calb", label = "pH spectro no calibration", color =  "black", pointShape = "triangle", strokeWidth = 0, pointSize = 6 ) %>%
dyEvent(as.POSIXct("2017-08-24 12:00:00", tz="UTC"), "seaFET #1005 installed", labelLoc = "bottom", color="red") %>%
  dyEvent(as.POSIXct("2018-02-02 09:00:00", tz="UTC"), "seaFET settings changed to every 1H", labelLoc = "bottom", color="red") %>% 
  dyEvent(as.POSIXct("2018-04-17 12:00:00", tz="UTC"), "seaFET #0007 installed + continuous mode set", labelLoc = "bottom", color="red") %>%
    dyEvent(as.POSIXct("2019-09-02 12:00:00", tz="UTC"), "seaFET #0007 out of water for cleaning", labelLoc = "bottom", color="red") %>%
  dyShading(from =  sfinterv[1], to= sfinterv[2], color = "#CCCCCC") %>%
  dyShading(from =  sfinterv[2], to= sfinterv[3], color = "#E0E0E0") %>%
  dyShading(from = sfinterv[3], to= Sys.Date(), color = "#E0E0E0") %>%
  dyLegend(show = "auto")   %>%
  dyAxis("y2",valueRange = c(-2, 2)) %>%
  dyHighlight(highlightCircleSize = 8,highlightSeriesBackgroundAlpha = 0.5,hideOnMouseOut = TRUE) %>%
  dyOptions(drawGrid = TRUE, drawPoints = TRUE, pointSize = c(1,8,1),useDataTimezone = TRUE, digitsAfterDecimal=4) %>%
  dyRangeSelector(height = 30, dateWindow= NULL)

#RANGE of pH
d_all<- d_all%>%
  dplyr::mutate(year = year(datetime))
range <- tbl_df(
  d_all %>%
  group_by(date=as.Date(datetime), 
           month=month(datetime, label=TRUE), 
           year=year(datetime)) %>%
  summarize(pH_range = max(pHint_tot_sf) - min(pHint_tot_sf))
)
quant <- quantile(range$pH_range, probs =c(2.5/100, 97.5/100), na.rm =TRUE)

#Per month
range_month <-
  range %>%
  group_by(Month = month(date, label=TRUE, abbr=FALSE)) %>%
  summarize(pH_range_month = mean(pH_range, na.rm=TRUE))

size_point <- 0.4 #size of data points
p <-  
  ggplot(data=range, aes(x = as.factor(month), y = pH_range)) +
  geom_violin(fill='lightblue', alpha=0.5, size=0.2, trim = TRUE) +
  geom_boxplot(notch=TRUE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(y =expression(paste(pH[T], " range seaFET")), x=NULL, title=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank())
print(p) # plot_range_sf
#ggsave(paste0("figures/plot_range_sf.png"), width= 30, height=10, plot_range_sf, units="cm")
```

>>>>

```{r Read, cleanup and prepare ferrybox data, echo=FALSE, warning=FALSE, message=FALSE}
d_hour <-
  readRDS(file = paste0(
    path,
    "fb_awipev-co2_server/ny-alesund/data/NRT_data/nydata_hour.rds"
  ))
d_hour <- d_hour %>%
  dplyr::rename(
    pco2_corr = pco2_corr_filtered,
    temp_dur = temp_dur_filtered,
    pco2_raw = pco2_raw_filtered,
    sal_fb = sal_fb_filtered,
    temp_insitu_11m = temp_insitu_11m_filtered,
    temp_fb = temp_fb_filtered,
    ph_dur = ph_dur_filtered
  ) %>% 
  arrange(datetime)
d_all <- as_tibble(d_hour) #d_all will be the merger
d_all <- d_all %>%
  dplyr::mutate_all(function(x) ifelse(is.nan(x), NA, x)) %>% # replace all NaN by NA
  dplyr::rowwise() %>% 
  dplyr::mutate(n_depth_ctd = sum(is.na( # create new variable to identify the hours for which there are CTD data at more than 1 depth 
  c(
    sal_insitu_ctd_1m,
    sal_insitu_ctd_3m,
    sal_insitu_ctd_5m,
    sal_insitu_ctd_7m,
    sal_insitu_ctd_9m
  )
), na.rm = TRUE)) %>% 
  ungroup()
#restore datetime, date which are mangled by the mutates above, arrange ascending
d_all <- d_all %>% 
  dplyr::mutate(datetime = d_hour$datetime, 
                       date = d_hour$date,
                       month = month(d_hour$datetime))
```

@@

# seaFET pH 
discrete2merge <- discrete %>%
  dplyr::select(closest_datetime, ph_s_sf_temp_insi) %>%
  dplyr::filter(!is.na(closest_datetime)) %>%
  rename(datetime=closest_datetime)
d_all <- full_join(d_all, discrete2merge, by="datetime")
d_all <- d_all %>% 
  dplyr::mutate(ph_s_sf_temp_insi = ifelse(n_depth_ctd == 4, ph_s_sf_temp_insi, NA)) # eliminate ph spec data when the profiler was at several depth during the same hour
  #dplyr::mutate(ph_s_sf_temp_insi = ifelse(!is.na(temp_insitu_ctd_7m) |
  #                                           !is.na(temp_insitu_ctd_9m), 
   #                                        ph_s_sf_temp_insi, NA)) # only use ph spec data when profiler was below 6 m

@@

# bind new calibrated pH data to the seafet data set.
names(seafet_calib) <- c("pHint_tot_sf", "pHext_tot_sf")
d_all <- dplyr::bind_cols(d_all, seafet_calib )

# pH seaFET Final
# add despike with n=2 k = 5762 to the final ph corrected data
# despike me semble inutile, faire plutôt un simple des outliers qui, vérification faite, n'est même pas nécessaire
# Il faut par contre créér une nouvelle variable avec pH deeper than 8 m, called *_9m for consistency with salinity and temp
d_all <- d_all %>%
  #dplyr::mutate(pHint_tot_sf = despike(d_all$pHint_tot_sf, reference= "median", n=2, k=5761, replace="NA")) %>%
  dplyr::mutate(pHint_tot_sf = seafet_calib$pHint_tot_sf) %>%
  dplyr::mutate(pHint_tot_sf_9m = ifelse(pressure_insitu_ctd > 8, pHint_tot_sf, NA),
                pHint_tot_sf_6m = ifelse(pressure_insitu_ctd > 6, pHint_tot_sf, NA))

@@

p <- ggplot(d_all, aes(x=datetime, y= pHint_tot_sf_9m))+
  geom_point(size=1, alpha=0.5, na.rm=TRUE)
p

seafet_depth_xts <- dplyr::select(d_all, datetime, ph_s_sf_temp_insi, pHint_tot_sf_6m, pHint_tot_sf_9m)
seafet_depth_xts <- as.xts(seafet_depth_xts, order.by = d_all$datetime)
dygraph(seafet_depth_xts, group = "depth", main="pH seaFET", ylab="pHT") %>%
dySeries("ph_s_sf_temp_insi", label = "pH spectro", color =  "red", strokeWidth = 0, pointSize = 4 ) %>%
dySeries("pHint_tot_sf_9m", label = "pHT 9m", color =  RColorBrewer::brewer.pal(3, "Dark2")[1], strokeWidth = 0, pointSize = 1) %>%
dySeries("pHint_tot_sf_6m", label = "pHT 6m", color =  RColorBrewer::brewer.pal(3, "Dark2")[2], strokeWidth = 0, pointSize = 1) %>%
dyEvent(as.POSIXct("2017-08-24 12:00:00", tz="UTC"), "seaFET #1005 installed", labelLoc = "bottom", color="red") %>%
  dyEvent(as.POSIXct("2018-02-02 09:00:00", tz="UTC"), "seaFET settings changed to every 1H", labelLoc = "bottom", color="red") %>% 
  dyEvent(as.POSIXct("2018-04-17 12:00:00", tz="UTC"), "seaFET #0007 installed + continuous mode set", labelLoc = "bottom", color="red") %>%
    dyEvent(as.POSIXct("2019-09-02 12:00:00", tz="UTC"), "seaFET #0007 out of water for cleaning", labelLoc = "bottom", color="red") %>%
  dyShading(from =  sfinterv[1], to= sfinterv[2], color = "#CCCCCC") %>%
  dyShading(from =  sfinterv[2], to= sfinterv[3], color = "#E0E0E0") %>%
  dyShading(from = sfinterv[3], to= Sys.Date(), color = "#E0E0E0") %>%
  dyLegend(show = "auto")   %>%
  dyAxis("y2",valueRange = c(-2, 2)) %>%
  dyHighlight(highlightCircleSize = 8,highlightSeriesBackgroundAlpha = 0.5,hideOnMouseOut = TRUE) %>%
  dyOptions(drawGrid = TRUE, drawPoints = TRUE, pointSize = c(1,8,1),useDataTimezone = TRUE, digitsAfterDecimal=4) %>%
  dyRangeSelector(height = 30, dateWindow= NULL)

#RANGE of pH
d_all<- d_all%>%
  dplyr::mutate(year = year(datetime))
range <- d_all %>%
  group_by(date=as.Date(datetime), 
           month=month(datetime, label=TRUE), 
           year=year(datetime)) %>%
  summarize(pH_range_9m = max(pHint_tot_sf_9m) - min(pHint_tot_sf_9m),
            pH_range_6m = max(pHint_tot_sf_6m) - min(pHint_tot_sf_6m))

quant <- quantile(range$pH_range_9m, probs =c(2.5/100, 97.5/100), na.rm =TRUE)
print(quant)
quant <- quantile(range$pH_range_6m, probs =c(2.5/100, 97.5/100), na.rm =TRUE)
print(quant)

#Per month
range_month <-
  range %>%
  group_by(Month = month(date, label=TRUE, abbr=FALSE)) %>%
  summarize(pH_range_month6m = mean(pH_range_6m, na.rm=TRUE),
            pH_range_month_9m = mean(pH_range_9m, na.rm=TRUE)) %>%
  ungroup()

size_point <- 0.4 #size of data points
p <- range_month %>%
  pivot_longer(-Month, names_to = "depth", values_to = "pH_range") %>%
  ggplot() +
  geom_violin(aes(x = as.factor(Month), y = pH_range), fill='lightblue', alpha=0.5, size=0.2, trim = TRUE) +
  geom_boxplot(notch=TRUE, size=0.2, fill="grey50", outlier.color = "blue", outlier.size = 0.3) +
  labs(y =expression(paste(pH[T], " range seaFET")), x=NULL, title=NULL) +
  Mytheme() +
  theme(panel.grid.major.x = element_blank())
print(p) # plot_range_sf
#ggsave(paste0("figures/plot_range_sf.png"), width= 30, height=10, plot_range_sf, units="cm")
```

### **seaFET pH profiles**
